---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r}
library(lattice)
library(ggplot2)
library(nnet)
```

## Loading and preprocessing the data
```{r}
temp <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp,method="curl")
data <- read.csv(unz(temp, "activity.csv"), header=TRUE)
unlink(temp)
```
The required histogram of the number of steps is as follows: 
````{r}
stepsperday<-aggregate(steps ~ date, data, sum)
hist(stepsperday$steps)
```

## What is mean total number of steps taken per day?
```{r}

mean(stepsperday$steps, na.rm=TRUE)
median(stepsperday$steps,na.rm=TRUE)
```
## What is the average daily activity pattern?
The "time series plot" of the 5-minute interval and the average number of steps taken, averaged across all days.  
```{r}
aggdata<-aggregate(steps ~ interval, data, mean)
xyplot(aggdata$steps~aggdata$interval,type="l")
```
  
The maximum value of steps is found at the interval 835: 
```{r}
index<-which.is.max(aggdata$steps)
aggdata[index,]
```

## Imputing missing values
The number of complete cases is `r sum(complete.cases(data))`; the cases with missing data is therefore `r 17568-sum(complete.cases(data))`.
I create a new dataframe, for imputing the missing data. 
```{r}
dataNoNa<-data
a<-which(is.na(data$steps))
for (index in 1:length(a)) {
        lineaggr<-which(aggdata$interval==dataNoNa$interval[a[index]])
        dataNoNa$steps[a[index]]<-aggdata$steps[lineaggr]        
}
```
Now, the histogram, mean and median of steps are the following: 
````{r}

stepsperday2<-aggregate(steps ~ date, dataNoNa, sum)
hist(stepsperday2$steps)
mean(stepsperday2$steps, na.rm=TRUE)
median(stepsperday2$steps,na.rm=TRUE)
```
The impact of imputing the missing data on these measures is very small; only the median is changed, and only slightly. 
## Are there differences in activity patterns between weekdays and weekends?
A new factor variable is created, indicating whether it is the weekend or not. 
```{r}
dates<-as.Date.factor(dataNoNa$date)
weekdays<-weekdays(dates)
weekend<-which(weekdays=="Samedi"|weekdays=="Dimanche") ##These are in French automatically; I have not fought this battle. 
week<-which(weekdays!="Samedi"&weekdays!="Dimanche")
dataNoNa$weekornot<-c(dataNoNa$date)
dataNoNa$weekornot[week]<-0
dataNoNa$weekornot[weekend]<-1
dataNoNa$weekornot<-factor(dataNoNa$weekornot,levels=c(0,1),labels<-c("weekdays","weekend"))
```
Now, the plot is created. 
```{r}
aggdata2<-aggregate(dataNoNa$steps,list(interval=dataNoNa$interval,weekornot=dataNoNa$weekornot), mean)
qplot(interval,x,data=aggdata2,facets=weekornot~.,geom="line",ylab="steps")
```



